---
title: "Problem Set 4"
author: "Zhengting (Johnathan) He"
date: "2021/12/10"
output: html_document
---

```{r "setup", include = FALSE}
require("knitr")
opts_knit$set(root.dir = "D:/OneDrive - Johns Hopkins/Course/140.622.81 - Statistical Methods in Public Health II/Problem Set/jhsphbiostat622-assignment/Problem Set 4")
```


```{r, message = FALSE}
setwd("D:/OneDrive - Johns Hopkins/Course/140.622.81 - Statistical Methods in Public Health II/Problem Set/jhsphbiostat622-assignment/Problem Set 4")

require(tidyverse)
require(tidymodels)
require(ResourceSelection)

nepalibf <- read_csv("./data/nepalibf.csv")
```


**1. Compare boxplots for the continuous covariates (X’s) vs breastfeeding status:**


```{r}
nepalibf %>% 
    mutate(bf = recode_factor(bf, `0` = "no", `1` = "yes")) %>% 
    drop_na() %>% 
    ggplot(aes(x = bf, y = age_chld)) +
    stat_boxplot(geom = "errorbar", width = 0.5) +
    geom_boxplot(aes(fill = bf)) +
    xlab("Breast feeding") +
    ylab("Age of child (months)") +
    scale_x_discrete(labels = c("No", "Yes")) +
    scale_fill_manual(values = c("#439fd3", "#f6b128"),
                      guide = "none") +
    theme_bw()
```


```{r}
nepalibf %>% 
    mutate(bf = recode_factor(bf, `0` = "no", `1` = "yes")) %>% 
    drop_na() %>% 
    ggplot(aes(x = bf, y = age_mom)) +
    stat_boxplot(geom = "errorbar", width = 0.5) +
    geom_boxplot(aes(fill = bf)) +
    xlab("Breast feeding") +
    ylab("Age of mom (years)") +
    scale_x_discrete(labels = c("No", "Yes")) +
    scale_fill_manual(values = c("#439fd3", "#f6b128"),
                      guide = "none") +
    theme_bw()
```


```{r}
nepalibf %>% 
    mutate(bf = recode_factor(bf, `0` = "no", `1` = "yes")) %>% 
    drop_na() %>% 
    ggplot(aes(x = bf, y = parity)) +
    stat_boxplot(geom = "errorbar", width = 0.5) +
    geom_boxplot(aes(fill = bf)) +
    xlab("Breast feeding") +
    scale_x_discrete(labels = c("No", "Yes")) +
    scale_fill_manual(values = c("#439fd3", "#f6b128"),
                      guide = "none") +
    theme_bw()
```


**2. Tabulate the proportion breastfed by gender:**


```{r}
CT = xtabs(~ sex_chld + bf, data = nepalibf)
addmargins(CT)
prop.table(CT, margin = 1)
addmargins(prop.table(CT, margin = 1), margin = 2)
```


**3. Briefly describe your findings**


---

Breast feeding probabilities were quite different across age of child based on crude estimates: children who were breast fed were younger (median [25th percentile, 75th percentile]: 50 [38, 59]) then those who were not breast fed (median [25th percentile, 75th percentile]: 19 [12, 28]), indicating age of child may be associated or be able to explain probabilities of breast feeding.

Breast feeding probabilities were not quite difference across age of mom, number of live births a mother has had, or sex of child, based on crude estimates. It is noticable that the crude estimates may be confounded by other variables.

---


**4. Estimate the prevalence of breast feeding as a function of child’s age (centered at the mean age) and gender using logistic regression:**


```{r}
nepalibf <- nepalibf %>%
    mutate(agechldc = age_chld - mean(age_chld))

model1 <- logistic_reg() %>%
    set_engine("glm") %>%
    set_mode("classification") %>%
    fit(as.factor(bf) ~ sex_chld + agechldc, data = nepalibf)
model1 %>% tidy() %>% kable()
model1 %>% tidy(exponentiate = TRUE) %>% kable()
exp(confint(model1 %>% extract_fit_engine()))
```


**a) Define the variables and specify the logistic regression model.**


---

+ Dependent variable: $Y$ = breast feed

+ Independent variable:

  - $X_1$ = child's gender
  
  - $X_2$ = age of child (months) - 37.338 months (center at 37.338 months)
  
+ Logistic regression model:

$$
log \, \frac{P(Y=1)}{1-P(Y=1)} = b_0 + b_1*X_1 + b_2*X_2
$$

---


**b) Interpret the coefficients $\hat{\beta}$; also interpret $e^{\hat{\beta}}$**


---

+ $\hat{\beta_0} = -0.92$: log odds of breast feeding among boys age at 37.338 months.

+ $\hat{\beta_1} = -0.31$: log odds ratio of breast feeding comparing girls to boys at the same age.

+ $\hat{\beta_2} = -0.19$: log odds of breast feeding decrease by 19% with 1 month increase in age among children of the same gender. (Log odds ratio of breast feeding comparing children at a certain age + 1 month versus children at a certain age is -0.19.)

+ $e^{\hat{\beta_0}} = 0.40$: odds of breast feeding among boys age at 37.338 months.

+ $e^{\hat{\beta_1}} = 0.73$: odds ratio of breast feeding comparing girls to boys at the same age.

+ $e^{\hat{\beta_2}} = 0.83$: odds of breast feeding decrease by 17% with 1 month increase in age among children of the same gender. (Odds ratio of breast feeding comparing children at a certain age + 1 month versus children at a certain age is 0.83.)

---


**5. Test the hypothesis that the association of breast feeding prevalence and child’s age is different for boys and girls (interaction model).**


```{r}
modelE <- logistic_reg() %>%
    set_engine("glm") %>%
    set_mode("classification") %>%
    fit(as.factor(bf) ~ agechldc + sex_chld + sex_chld:agechldc, data = nepalibf)
modelE %>% tidy() %>% kable()
```


**a) Is the evidence of a different age-prevalence relationship for boys and girls weak, moderate, or strong?**


---

According to the output of Model E, the odds of breast feeding decrease by 17% with 1 month increase in age among boys, and the odds of breast feeding decrease by 195 with 1 month increase in age among girls. Since the z-statistic associated with the coefficient of the interaction term is -1.007, and the p-value is 0.31, the coefficient of the interaction term is not statistically significant under $\alpha = 0.05$, indicating that the evidence of a different age-prevalence relationship for boys and girls is weak.

---


**b) Interpret the coefficients $\hat{\beta_3}$ for the interaction and $e^{\hat{\beta_3}}$.**


---

+ $\hat{\beta_3} = -0.04$: difference in the log odds ratio (log {ratio of odds ratio}) of breast feeding comparing children at a certain age + 1 month to a certain age, among boys versus girls.

+ $e^{\hat{\beta_3}} = 0.96$: ratio of odds ratio of breast feeding comparing children at a certain age + 1 month to a certain age, among boys and girls.

+ Since the z-statistic associated with the coefficient of the interaction term is -1.007, and the p-value is 0.31, the coefficient of the interaction term is not statistically significant under $\alpha = 0.05$, indicating that the evidence of a different age-prevalence relationship for boys and girls is weak.

+ The results are consistent using Likelihood Ratio Test below.

---


```{r}
anova(model1 %>% extract_fit_engine(), modelE %>% extract_fit_engine(), test = "LRT") %>% tidy() %>% kable()
```


**6. Display a graph of the estimated breast feeding prevalence -vs- child’s age with separate curves for boys and girls:**


```{r}
nepalibf <- nepalibf %>%
    na.omit() %>% # Remove observations with missing data
    mutate(sex_chld = recode_factor(sex_chld, `0` = "Male", `1` = "Female")) # Factor sex_chld

modelD <- logistic_reg() %>%
    set_engine("glm") %>%
    set_mode("classification") %>%
    fit(as.factor(bf) ~ sex_chld + age_chld, data = nepalibf)

nepalibf <- nepalibf %>% mutate(phat = predict(modelD %>% extract_fit_engine(), type = "response"))

nepalibf %>% 
    ggplot(aes(x = age_chld, y = phat, color = sex_chld, shape = sex_chld)) +
    geom_point() +
    scale_color_manual(values = c("#439fd3", "#f6b128")) +
    theme_bw()
```


**Which gender has the highest breastfeeding prevalence and by how much higher is it? (Base this on your interpretation of the graph only.)**


---

According to the graph, male has a higher predicted prevalence of breastfeeding than female. The difference of breastfeeding prevalence between male and female varies by child age, with lowest difference in the two-side tails of child age distribution, and highest difference in the middle of child age distribution. The mean difference is around 2%, with highest difference around 8%.

---


**7. Carry out a Hosmer-Lemeshow goodness-of- fit test for the model that includes gender and child’s age. Divide the predicted probabilities into 10 roughly equal groups for calculation of the chi-square comparing observed and expected events across the groups:**


```{r}
hoslem.test(nepalibf$bf, nepalibf$phat, g = 10)
hoslem.test(nepalibf$bf, nepalibf$phat)$observed
hoslem.test(nepalibf$bf, nepalibf$phat)$expected
```


**Interpret the Hosmer-Lemeshow P-value to determine whether there is evidence of poor fit for this model.**


---

According to the Hosmer-Lemeshow test, the test statistic $\chi^2$ = 6.0898, and p-value = 0.6405 under df = 8. Under $\alpha$ = 0.05, we fail to reject $H_0$ and conclude the model fit is good.

---


**8. Summarize in one page, as if for a journal article, the findings of your analysis of the association between the decision to breastfeed and characteristics of the child.**


---

According to the outputs above, there is an association between child's age, gender and breastfeeding. The odds ratio of breastfeeding decrease by 18% with 1 month increase in child's age, after adjusting for gender; and the association is statistically significant under $\alpha$ = 0.05. The odds ratio of breastfeeding comparing girls to boys is 0.73, after adjusting for child's age; however, the association is not statistically significant under $\alpha$ = 0.05. Gender does not statistical significantly modifies the association between child's age and breastfeeding. Goodness of fit test shows the main logistic regression model fit is good.

---

