---
title: "Problem Set 2"
author: "Zhengting (Johnathan) He"
date: "2021/11/13"
output: html_document
---


```{r "setup", include = FALSE}
require("knitr")
opts_knit$set(root.dir = "D:/OneDrive - Johns Hopkins/Course/140.622.81 - Statistical Methods in Public Health II/Problem Set/jhsphbiostat622-assignment/Problem Set 2")
```


```{r, message = FALSE}
setwd("D:/OneDrive - Johns Hopkins/Course/140.622.81 - Statistical Methods in Public Health II/Problem Set/jhsphbiostat622-assignment/Problem Set 2")

require(tidyverse)
require(tidymodels)
```


# I. Multiple linear regression to compare “growth” for male and female infants (Analysis of Covariance or ANCOVA)


**0. Using the Nepal Children’s Anthropometry Dataset for children ages 0-60 months, omit children with ages greater than 12 months. Also omit all children for whom height, weight or arm circumference is missing.**


```{r, message = FALSE}
nepalA <- read_csv("./data/nepal_anthro.csv")
```


```{r}
nepalData <- nepalA %>%   ### store this dataset under a new name
   filter(age <= 12) %>%   ### keep only children with age <= 12
   drop_na("height", "weight", "armcirc")   ### remove NAs
```


**1. Make a scatter-plot of weight against age discriminating boy and girl infants by the plotting color. Use a small plotting symbol (e.g. dot) and jitter the data so all points can be seen as a cloud. Recode sex as a 0/1 factor variable named gender with two levels, “Male” and “Female”, where “Male” is the reference group (the 0).**


```{r}
nepalData <- nepalData %>% 
    mutate(gender = recode_factor(sex, `1`="Male", `2`="Female"))
    ### the first level you designate will be the reference!
```


```{r}
theme_622 <- function() {
    theme(
        panel.background = element_rect(fill = 'transparent', color = 'gray'),
        panel.border = element_rect(fill = 'transparent', color = 'transparent'),
        panel.grid = element_line(colour = "grey92"),
        panel.grid.minor = element_line(size = rel(0.5)),
        axis.title.x = element_text(size = 16, face = "bold", color = "black"),
        axis.title.y = element_text(size = 16, face = "bold", color = "black"),
        axis.text = element_text(size = 12, color = "black"),
        legend.title = element_text(size = 12, face = "bold", color = "black"),
        legend.text = element_text(size = 12, color = "black"),
        legend.position = "bottom"
    )
}
```


```{r}
qplot(x = jitter(age), y = weight, color = gender, shape = gender, 
      data = nepalData, xlab = "Age in months", ylab = "Weight in kg") +
    theme_622()
```


**2. Fit a simple ANCOVA model (Model A) that includes gender, age and gender\*age (gender 
by age interaction). Obtain the fitted values from this model and add them to the graph using 
the same colors for the boy and girl predicted values as for the original data.**


```{r}
model1 <- linear_reg() %>%
    set_engine("lm") %>%
    fit(weight ~ age + gender, data = nepalData) %>%
    extract_fit_engine()
```


```{r}
summary(model1) %>% tidy() %>% kable()
```


```{r}
anova(model1) %>% tidy() %>% kable()
```


```{r}
model2 <- linear_reg() %>%
    set_engine("lm") %>%
    fit(weight ~ age*gender, data = nepalData) %>%
    extract_fit_engine()
```


```{r}
summary(model2) %>% tidy() %>% kable()
```


```{r}
anova(model2) %>% tidy() %>% kable()
```


```{r}
qplot(x = jitter(age), y = weight, color = gender, shape = gender,
      data = nepalData, xlab = "Age in months", ylab = "Weight in kg") +
    geom_line(aes(x = age, y=model2$fitted.values, color=gender)) +
    theme_622()
```


**3. Interpret each of the coefficients in this model. Do the data provide evidence that birth 
weights are different for boys and girls; growth rates? Be quantitative, citing the appropriate 
estimates and standard errors (or tests) in your answers.**


---

+ $\hat{\beta_0}$ = 3.89: the expected mean weights for boys is 3.89kg at birth (age = 0 months).

+ $\hat{\beta_1}$ = 0.34: the mean weights is expected to increase 0.34kg per 1 month increase in age, among boys.

+ $\hat{\beta_2}$ = -0.42: the expected difference in mean weights for girls compared with boys at birth (age = 0 months).

+ $\hat{\beta_3}$ = -0.02: the expected difference in mean weights changes per 1 month increase in age for girls compared with boys.

+ Birth weights are statistically significantly different for boys and girls. This is supported by the evidence that t-value for $\hat{\beta_2}$ is -2.948, and the p-value is 0.00328, and by the F-value = 52.2374, and p-value = $1.055*10^{-12}$, thus under $\alpha = 0.05$, reject $H_0$ and conclude that birth weights are statistically significantly difference for boys and girls.

+ Weight growth rate is statistically significantly different than 0. This is supported by the evidence that t-value for $\hat{\beta_1}$ is 24.488, and the p-value < $2*10^{-16}$, and by the F-value = 1154.8236, and p-value < $2*10^{-16}$, thus under $\alpha = 0.05$, reject $H_0$ and conclude that weight growth rate is statistically significantly different than 0.

---


**4. Calculate the residuals from this model and plot them against age using the same colors for 
boys and girls as were used for the raw data. Comment whether this model, that assumes 
growth is linear in the first year, is adequate.**


```{r}
qplot(y = model2$residuals, x = jitter(age), color = gender, shape = gender, 
      data = nepalData, ylab = "Residuals", xlab = "Age in months") +
    theme_622()
```

---

Based on the residual plot, the assumption of this model that growth is linear in the first year may not hold. According to the plot, throughout the range of $X$, the residual does not appear to be scatter equally around the value of $0$, but has a curved trend, both for boys and girls. The trend can be showed clearer using lowess regression.

---


```{r}
qplot(y = model2$residuals, x = jitter(age), color = gender, shape = gender, 
      data = nepalData, ylab = "Residuals", xlab = "Age in months") +
    geom_smooth(method = "loess", formula = y ~ x) +
    theme_622()
```


**5. Add a linear spline term with a break at 4 months and its interaction with gender to the model
above. Call the new model B. Calculate and plot the fitted values from Model B against age. 
Interpret each of the coefficients in Model B.**


```{r}
nepalData <- nepalData %>%
    mutate(agesp = ifelse(age > 4, age-4, 0))
```


```{r}
model3 <- linear_reg() %>%
    set_engine("lm") %>%
    fit(weight ~ age*gender + agesp*gender, data=nepalData) %>%
    extract_fit_engine()
```


```{r}
summary(model3) %>% tidy() %>% kable()
```


```{r}
confint(model3)
```


```{r}
qplot(x = jitter(age), y = weight, color = gender, shape = gender, 
      data = nepalData, xlab = "Age in months", ylab = "Weight in kg") +
    geom_line(aes(x = age, y=model3$fitted.values, color=gender), size = 1) +
    theme_622()
```


---

$\hat{weight} = b_0 + b_1*age + b_2*gender + b_3*(age-4)^+ + b_4*age:gender + b_5*(age-4)^+:gender$

1. When gender = male, age $\leq$ 4: $\hat{weight} = b_0 + b_1*age$

2. When gender = male, age > 4: $\hat{weight} = (b_0 - 4*b_3) + (b_1 + b_3)*age$

3. When gender = female, age $\leq$ 4: $\hat{weight} = (b_0 + b_2) + (b_1 + b_4)*age$

4. When gender = female, age > 4: $\hat{weight} = (b_0 + b_2 - 4*(b_3 + b_5)) + (b_1 + b_3 + b_4 + b_5)*age$

+ $\hat{\beta_0}$ = 2.79: the expected mean weights for boys is 2.79kg at birth (age = 0 months).

+ $\hat{\beta_1}$ = 0.79: the mean weights is expected to increase 0.79kg per 1 month increase in age, when age $\leq$ 4 in boys.

+ $\hat{\beta_2}$ = -0.21: the expected difference in mean weights for girls compared with boys at birth (age = 0 months).

+ $\hat{\beta_3}$ = -0.58: the expected difference in the change of mean weights per 1 month increase in age when age > 4, compared with when age $\leq$ 4, in boys.

+ $\hat{\beta_4}$ = -0.09: the expected difference in mean weights changes per 1 month increase in age for girls compared with boys, when age $\leq$ 4.

+ $\hat{\beta_5}$ = 0.09: the expected differences - in the difference in mean weights changes per 1 month increase in age for girls compared with boys - when age > 4 compared with when age $\leq$ 4.

---


**6. Test the null hypothesis that growth is linear over the first twelve months by testing whether
the coefficients of the two new terms in Model B are both zero.**


```{r}
anova(model2, model3)
```


**Note: here’s another way to do it!**


```{r}
survey::regTermTest(model3, ~ agesp + agesp:gender)
```


---

According to the ANOVA test comparing Model A and Model B, F-statistic = 71.74 on $df_{2,892}$, and corresponding p-value $\leq 2.22*10^{-16}$. Thus reject $H_0$ under $\alpha = 0.05$ and conclude that growth is not linear over the first twelve months.

---


**7. Calculate the residuals from this model and plot them against age. Comment on any 
assumptions of the linear regression that still appear to be violated.**


```{r}
qplot(y = model3$residuals, x = jitter(age), data = nepalData,
      ylab = "Residuals", xlab = "Age in months") +
    geom_hline(yintercept = 0, color = "red") +
    theme_622()
```


---

According to the residual plot of residual over age, the residual seems to scatterly and randomly distributed around $0$ over the range of $X$, and no apparent trend identified, indicating that the linear assumption was not violated.

---


**8. Write a paragraph to summarize your findings regarding growth of boys and girls as if for a 
public health journal. Be quantitative (use numbers) and avoid statistical jargon.**


---

Comparing to birth weights of girls, birth weights of boys was approximately 0.21kg higher; however it is noticed that the difference is not statistically significant. Within 4 months after birth, weights of boys increase approximately 0.79kg with 1 month increase in age, and weights of girls increase approximately 0.70kg with 1 month increase in age. Gender may be an effect modifier on the association between age and weight growth of children; however, it is noticed that the effect modification is not statistically significant. After the age of 4 months, the weight growth of boys and girls was approximately 0.58kg and 0.49kg statistically significantly slower, respectively, compared with the first 4 months after birth. This result in a similar weight growth of approximately 0.20kg in both boys and girls per 1 month increase in age after the age of 4 months. The global F-statistic and p-value shows the significant of the linear relationship (at least one parameters $\leq$ 0), and $R^2$ value is 0.6336, indicates a moderate ability of predicting weight using the covarates of age, gender, age spline term, and interaction term of age and gender, interaction term of age spline term and gender.

---


# II. Modelling Non-linear Relationships with MLR


**1. Use the data set from Problem I.**


**2. Plot weight against age (use small plotting symbols, e.g. dots, and jitter the points so they can all be seen).**


```{r}
qplot(x = jitter(age), y = weight, data = nepalData,
      xlab = "Age in months", ylab = "Weight in kg", ylim = c(0, 12)) +
    theme_622()
```


**3. Calculate the sample mean weight for each month of age and add the estimated means for 
each month to the graph (with bold symbols and a connecting line) to highlight the trend.**


**To compute mean weight for each month of age will work because age takes on 13 unique 
values (the integers 0-12): if age was truly continuous, the following command would fail to 
work.**


```{r}
nepalData %>% group_by(age) %>% summarize(mean = mean(weight)) %>% kable()
```


**To add means to previous graph:**


```{r}
qplot(x = jitter(age), y = weight, data = nepalData,
    xlab = "Age in months", ylab = "Weight in kg", ylim = c(0, 12)) +
    stat_summary(aes(x = age, y = weight), fun.y = mean, geom = "line",
                 lwd = 2, color = "red") +
    theme_622()
```


**(Scientific digression: this pattern actually represents the convolution of two biological 
processes: growth and seasonality. For simplicity, below we will refer to it as “growth.”)**


**4. Use simple linear regression to regress weight on age and add the least squares line to the 
plot.**


```{r}
model4 <- linear_reg() %>%
    set_engine("lm") %>%
    fit(weight ~ age, data = nepalData) %>%
    extract_fit_engine()
```


```{r}
summary(model4) %>% tidy() %>% kable()
```


```{r}
confint(model4)
```


```{r}
qplot(x = jitter(age), y = weight, data = nepalData, xlab = "Age in months",  
      ylab = "Weight in kg", ylim = c(0, 12)) +
    geom_smooth(method = "lm") +
    theme_622()
```


**OR, a more general way to plot “fitted values” on a plot:**


```{r}
qplot(x = jitter(age), y = weight, data = nepalData, xlab = "Age in months", 
      ylab = "Weight in kg", ylim = c(0, 12)) +
    geom_line(aes(x = age, y = model4$fitted.values), color = "red", lwd = 2) +
    theme_622()
```


**5. Calculate the residuals from the simple linear regression above and plot them versus age.
Add a smooth function using lowess smoothing. Comment in one sentence on the adequacy 
of a linear assumption for “growth”.**


```{r}
qplot(y = model4$residuals, x = jitter(age), data = nepalData, ylab = "Residuals",  
      xlab = "Age in months") + 
    geom_smooth(method = "loess", se = FALSE) +   # loess smoother
    geom_hline(yintercept = 0, color = "red") +   # horizontal y=0 line
    theme_622()
```


---

Based on the residual plot, the linear assumption for growth may not hold. According to the plot, throughout the range of $X$, the residual does not appear to be scatter equally around the value of $0$, but has a curved trend — residuals tends to be scattering below $0$ around 0~2 months and 10~12 months, while scattering above $0$ in 2~10 months.

---


**6. Use linear regression to regress weight onto the monthly mean ages.**


```{r}
model5 <- linear_reg() %>%
    set_engine("lm") %>%
    fit(weight ~ as.factor(age), data = nepalData) %>%
    extract_fit_engine()
```


```{r}
summary(model5) %>% tidy() %>% kable()
```


```{r}
confint(model5)
```


**7. Linear splines:**


+ **a. Create three new variables: **

age_sp1 = $(age – 2)^+ = age \ – 2$ if age > 2, 0 if not

age_sp2 = $(age - 4)^+ = age \ – 4$ if age > 4, 0 if not

age_sp3 = $(age - 6)^+ = age \ – 6$ if age > 6, 0 if not


```{r}
nepalData = nepalData %>%
    mutate(age_sp1 = ifelse(age > 2, age - 2, 0)) %>%
    mutate(age_sp2 = ifelse(age > 4, age - 4, 0)) %>%
    mutate(age_sp3 = ifelse(age > 6, age - 6, 0))
```


+ **b. Regress weight on age, age_sp1, age_sp2 and age_sp3.**


```{r}
model6 <- linear_reg() %>%
    set_engine("lm") %>%
    fit(weight ~ age + age_sp1 + age_sp2 + age_sp3, data = nepalData) %>%
    extract_fit_engine()
```


```{r}
summary(model6) %>% tidy() %>% kable()
```


```{r}
confint(model6)
```


+ **c. Get the predicted values from this regression. Add these to the plot of weight against age 
from step 1.**


```{r}
qplot(x = jitter(age), y = weight, data = nepalData, xlab = "Age in months",  
      ylab = "Weight in kg", ylim = c(0, 12)) +
    geom_line(aes(x = age, y = model6$fitted.values), color = "red", lwd = 2) +
    theme_622()
```


+ **d. Interpret the meaning of the coefficients for the three “linear spline” terms: age_sp1, 
age_sp2 and age_sp3.**


---

+ $\hat{\beta_0}$ = 2.63: the estimated weight at birth (age in months = 0)

+ $\hat{\beta_1}$ = 0.80: the estimated increase in weight per 1 month increase in age when age $\leq$ 2 months.

+ $\hat{\beta_2}$ = -0.16: difference between the estimated increase in weight per 1 month increase in age when age > 2 months and $\leq$ 4 months, compared with age $\leq$ 2 months.

+ $\hat{\beta_3}$ = -0.31: difference between the estimated increase in weight per 1 month increase in age when age > 4 months and $\leq$ 6 months, compared with age > 2 months and $\leq$ 4 months.

+ $\hat{\beta_4}$ = -0.16: difference between the estimated increase in weight per 1 month increase in age when age > 6 months, compared with age > 4 months and $\leq$ 6 months.

---


+ **e. Use an F-test with 3 degrees of freedom in the numerator to test the null hypothesis of a 
linear relationship against the alternative defined by the linear spline fit in the steps
above.**


```{r}
anova(model4, model6)
```


*OR*


```{r}
survey::regTermTest(model6, ~ age_sp1 + age_sp2 + age_sp3)
```


+ **f. Comment in a few sentences on the evidence from this analysis for or against a linear 
growth curve**


---

According to the ANOVA test comparing Model 4 and Model 6, F-statistic = 45.98 on $df_{3,893}$, and corresponding p-value $\leq 2.22*10^{-16}$. Thus reject $H_0$ under $\alpha = 0.05$ and conclude that growth is not linear over the first twelve months.

---


**8. Consider using your models to predict children’s weights. Complete the table below. 
PLEASE NOTE: AIC also can be calculated by hand using either of the formulas below.**


```{r}
AIC(model4, model5, model6) %>% kable()
```


**PLEASE NOTE: R explicitly count the variance as a parameter when calculating AIC for 
regression. So the calculations in R will have 3, 14, 6 as the degrees of freedom rather than 
the 2, 13, 5 shown in the table below. This just means that each AIC value calculated by R 
will be 2*1=2 units higher than the values when you calculate by hand according to this 
table!**


```{r}
AIC.table <- tibble(
    `Model (# parameters)` = c("Linear (2)", "Monthly means (13)", "Linear spline (5)"),
    `parameters` = c(2, 13, 5),
    `Residual sum of squares (RSS)` = c(anova(model4)$`Sum Sq`[length(anova(model4)$`Sum Sq`)],
                                        anova(model5)$`Sum Sq`[length(anova(model5)$`Sum Sq`)],
                                        anova(model6)$`Sum Sq`[length(anova(model6)$`Sum Sq`)]),
    `Residual mean square (MSE)` = c(anova(model4)$`Mean Sq`[length(anova(model4)$`Mean Sq`)],
                                     anova(model5)$`Mean Sq`[length(anova(model5)$`Mean Sq`)],
                                     anova(model6)$`Mean Sq`[length(anova(model6)$`Mean Sq`)]),
    `AIC.lm` = nrow(nepalData)*(log(2*pi*`Residual sum of squares (RSS)`/nrow(nepalData))+1) +
        2*`parameters`,
    `log-likelihood` = sapply(c(logLik(model4), logLik(model5), logLik(model6)), as.numeric),
    `AIC.glm` = -2*(`log-likelihood`) + 2*`parameters`
    ) %>%
    mutate(`parameters` = NULL)
```


```{r}
AIC.huxtable <- huxtable::as_hux(AIC.table) %>%
    huxtable::insert_row("", "Using the regress command", "", "", "Using the glm command", "") %>%
    huxtable::merge_cells(1, 2:4) %>%
    huxtable::merge_cells(1, 5:6) %>%
    huxtable::set_header_rows(1, TRUE) %>%
    huxtable::style_headers(bold = TRUE)
huxtable::number_format(AIC.huxtable) <- 2
AIC.huxtable
```


**9. Comment on the quality of each of the 3 models using the plots, the residual mean squares 
and AIC values, and your judgment about how children grow. Comment on what aspects of 
model quality each measure addresses.**

---

+ Quality of all 3 models:

  - Model 4: Model 4 is a simple linear regression using age to predict weight. The residual plot shows that linear assumption may not meet. $MSE$ = 1.21, which is the highest among all 3 models. $AIC$ = 2720.80, which is the highest among all 3 models. Based on these evidence, model 4 has the least quality for prediction among all 3 models.
  
  - Model 5: Model 4 is a multiple linear regression treating age as categorical data (indicator variable), and using linear regression to compare the mean weights across each age group. $MSE$ = 1.04, which is similar to the $MSE$ of model 6. $AIC$ = 2598.69, which is slightly higher than model 6. Based on these evidence, model 5 provides better quality for predicting weight compared with model 4; however, it does not use least parameters needed to predict weight, i.e., does not meet with parsimonious model criteria.
  
  - Model 6: Model 6 is a multiple linear regression including 3 spline terms. The scatter plot shows a relatively better quality fit for the data compared with model 4. $MSE$ = 1.05, which is similar to $MSE$ of model 5. $AIC$ = 2597.80, which is the lowest among all 3 models. Based on these evidence, model 6 has the highest quality for prediction among all 3 models.
  
+ Judgement on how children grow:

  - Based on the prediction models and their quality, model 6 has the highest quality for prediction, which shows that weight increases per each month increase in age. Children grow fastest within the first 2 months after baseline (0.799kg per 1 month increase in age); the weight increase rate decreases to 0.640kg per 1 month increase in age when age > 2 months; the weight increase rate decreases to 0.329kg per 1 month increase in age when age > 4 months; the weight increase rate decreases to 0.172kg per 1 month increase in age when age > 6 months. The changes in weight increase rate at each age cutoff point are statistically significant.

+ What aspects of model quality each measure addresses

  - Residual plot / scatter plot: whether assumptions of linear regression meets in the case of the model.
  
  - $MSE$: how much of the variability in weight cannot be explained by age.
  
  - $AIC$: approximation to the cross-validated prediction error (a criteria for determining the "best" model for prediction) 

---


**Create a script file to store your commands for this assignment. A script file makes it easy to re-run your commands later. It also makes it straightforward to reproduce your analysis at a late time. (See “Script files: keeping track of what you do” from the “R/RStudio: A Brief Introduction” document posted on Course Plus.)**


```
####################################### Problem Set 2 #######################################

# I. Multiple linear regression to compare “growth” for male and female infants (Analysis of
# Covariance or ANCOVA)

# Set up
require(tidyverse)
require(tidymodels)

setwd("D:/OneDrive - Johns Hopkins/Course/140.622.81 - Statistical Methods in Public Health II/Problem Set/jhsphbiostat622-assignment/Problem Set 2")

nepalA <- read_csv("./data/nepal_anthro.csv")

nepalData <- nepalA %>%   ### store this dataset under a new name
    filter(age <= 12) %>%   ### keep only children with age <= 12
    drop_na("height", "weight", "armcirc")   ### remove NAs


nepalData <- nepalData %>% 
    mutate(gender = recode_factor(sex, `1`="Male", `2`="Female"))
    ### the first level you designate will be the reference!


# Scatter plot
qplot(x = jitter(age), y = weight, color = gender, shape = gender, 
      data = nepalData, xlab = "Age in months", ylab = "Weight in kg") +
    theme(panel.background = element_rect(fill='transparent', color='gray'),
          panel.border = element_rect(fill='transparent', color='transparent'),
          panel.grid = element_line(colour = "grey92"), 
          panel.grid.minor = element_line(size = rel(0.5)),
          axis.title.x = element_text(size = 16, face = "bold", color = "black"),
          axis.title.y = element_text(size = 16, face = "bold", color = "black"),
          axis.text = element_text(size = 12, color = "black"),
          legend.title = element_text(size = 12, face = "bold", color = "black"),
          legend.text = element_text(size = 12, color = "black"),
          legend.position = "bottom")


# Linear regression
model1 <- linear_reg() %>%
    set_engine("lm") %>%
    fit(weight ~ age + gender, data = nepalData) %>%
    extract_fit_engine()
summary(model1)
anova(model1)

model2 <- linear_reg() %>%
    set_engine("lm") %>%
    fit(weight ~ age*gender, data = nepalData) %>%
    extract_fit_engine()
summary(model2)
anova(model2)


# Scatter plot with fitted line
qplot(x = jitter(age), y = weight, color = gender, shape = gender,
      data = nepalData, xlab = "Age in months", ylab = "Weight in kg") +
    geom_line(aes(x = age, y=model2$fitted.values, color=gender)) +
    theme(panel.background = element_rect(fill='transparent', color='gray'),
          panel.border = element_rect(fill='transparent', color='transparent'),
          panel.grid = element_line(colour = "grey92"), 
          panel.grid.minor = element_line(size = rel(0.5)),
          axis.title.x = element_text(size = 16, face = "bold", color = "black"),
          axis.title.y = element_text(size = 16, face = "bold", color = "black"),
          axis.text = element_text(size = 12, color = "black"),
          legend.title = element_text(size = 12, face = "bold", color = "black"),
          legend.text = element_text(size = 12, color = "black"),
          legend.position = "bottom")


# Residual plot
qplot(y = model2$residuals, x = jitter(age), color = gender, shape = gender, 
      data = nepalData, ylab = "Residuals", xlab = "Age in months") +
    theme(panel.background = element_rect(fill='transparent', color='gray'),
          panel.border = element_rect(fill='transparent', color='transparent'),
          panel.grid = element_line(colour = "grey92"), 
          panel.grid.minor = element_line(size = rel(0.5)),
          axis.title.x = element_text(size = 16, face = "bold", color = "black"),
          axis.title.y = element_text(size = 16, face = "bold", color = "black"),
          axis.text = element_text(size = 12, color = "black"),
          legend.title = element_text(size = 12, face = "bold", color = "black"),
          legend.text = element_text(size = 12, color = "black"),
          legend.position = "bottom")


qplot(y = model2$residuals, x = jitter(age), color = gender, shape = gender, 
      data = nepalData, ylab = "Residuals", xlab = "Age in months") +
    geom_smooth(method = "loess", formula = y ~ x) +
    theme(panel.background = element_rect(fill='transparent', color='gray'),
          panel.border = element_rect(fill='transparent', color='transparent'),
          panel.grid = element_line(colour = "grey92"), 
          panel.grid.minor = element_line(size = rel(0.5)),
          axis.title.x = element_text(size = 16, face = "bold", color = "black"),
          axis.title.y = element_text(size = 16, face = "bold", color = "black"),
          axis.text = element_text(size = 12, color = "black"),
          legend.title = element_text(size = 12, face = "bold", color = "black"),
          legend.text = element_text(size = 12, color = "black"),
          legend.position = "bottom")


# Linear spline
nepalData <- nepalData %>%
    mutate(agesp = ifelse(age > 4, age-4, 0))

model3 <- linear_reg() %>%
    set_engine("lm") %>%
    fit(weight ~ age*gender + agesp*gender, data=nepalData) %>%
    extract_fit_engine()
summary(model3) %>% tidy()
confint(model3)


# Testing on spline term and residual plot
anova(model2, model3)
survey::regTermTest(model3, ~ agesp + agesp:gender)

qplot(y = model3$residuals, x = jitter(age), data = nepalData,
      ylab = "Residuals", xlab = "Age in months") +
    geom_hline(yintercept = 0, color = "red") +
    theme(panel.background = element_rect(fill='transparent', color='gray'),
          panel.border = element_rect(fill='transparent', color='transparent'),
          panel.grid = element_line(colour = "grey92"), 
          panel.grid.minor = element_line(size = rel(0.5)),
          axis.title.x = element_text(size = 16, face = "bold", color = "black"),
          axis.title.y = element_text(size = 16, face = "bold", color = "black"),
          axis.text = element_text(size = 12, color = "black"),
          legend.title = element_text(size = 12, face = "bold", color = "black"),
          legend.text = element_text(size = 12, color = "black"),
          legend.position = "bottom")


# II. Modelling Non-linear Relationships with MLR

# Scatter plot
qplot(x = jitter(age), y = weight, data = nepalData,
      xlab = "Age in months", ylab = "Weight in kg", ylim = c(0, 12)) +
    theme(panel.background = element_rect(fill='transparent', color='gray'),
          panel.border = element_rect(fill='transparent', color='transparent'),
          panel.grid = element_line(colour = "grey92"), 
          panel.grid.minor = element_line(size = rel(0.5)),
          axis.title.x = element_text(size = 16, face = "bold", color = "black"),
          axis.title.y = element_text(size = 16, face = "bold", color = "black"),
          axis.text = element_text(size = 12, color = "black"),
          legend.title = element_text(size = 12, face = "bold", color = "black"),
          legend.text = element_text(size = 12, color = "black"),
          legend.position = "bottom")

nepalData %>% group_by(age) %>% summarize(mean = mean(weight))

qplot(x = jitter(age), y = weight, data = nepalData,
      xlab = "Age in months", ylab = "Weight in kg", ylim = c(0, 12)) +
    stat_summary(aes(x = age, y = weight), fun.y = mean, geom = "line",
                 lwd = 2, color = "red") +
    theme(panel.background = element_rect(fill='transparent', color='gray'),
          panel.border = element_rect(fill='transparent', color='transparent'),
          panel.grid = element_line(colour = "grey92"), 
          panel.grid.minor = element_line(size = rel(0.5)),
          axis.title.x = element_text(size = 16, face = "bold", color = "black"),
          axis.title.y = element_text(size = 16, face = "bold", color = "black"),
          axis.text = element_text(size = 12, color = "black"),
          legend.title = element_text(size = 12, face = "bold", color = "black"),
          legend.text = element_text(size = 12, color = "black"),
          legend.position = "bottom")


# Simple linear regression weight ~ age
model4 <- linear_reg() %>%
    set_engine("lm") %>%
    fit(weight ~ age, data = nepalData) %>%
    extract_fit_engine()
summary(model4) %>% tidy() %>% kable()
confint(model4)

qplot(x = jitter(age), y = weight, data = nepalData, xlab = "Age in months",  
      ylab = "Weight in kg", ylim = c(0, 12)) +
    geom_smooth(method = "lm") +
    theme_622()

qplot(x = jitter(age), y = weight, data = nepalData, xlab = "Age in months", 
      ylab = "Weight in kg", ylim = c(0, 12)) +
    geom_line(aes(x = age, y = model4$fitted.values), color = "red", lwd = 2) +
    theme_622()

qplot(y = model4$residuals, x = jitter(age), data = nepalData, ylab = "Residuals",  
      xlab = "Age in months") + 
    geom_smooth(method = "loess", se = FALSE) +   # loess smoother
    geom_hline(yintercept = 0, color = "red") +   # horizontal y=0 line
    theme_622()


# Linear regression weight ~ monthly mean age
model5 <- linear_reg() %>%
    set_engine("lm") %>%
    fit(weight ~ as.factor(age), data = nepalData) %>%
    extract_fit_engine()
summary(model5) %>% tidy()
confint(model5)


# Linear splines
nepalData = nepalData %>%
    mutate(age_sp1 = ifelse(age > 2, age - 2, 0)) %>%
    mutate(age_sp2 = ifelse(age > 4, age - 4, 0)) %>%
    mutate(age_sp3 = ifelse(age > 6, age - 6, 0))

model6 <- linear_reg() %>%
    set_engine("lm") %>%
    fit(weight ~ age + age_sp1 + age_sp2 + age_sp3, data = nepalData) %>%
    extract_fit_engine()
summary(model6) %>% tidy()
confint(model6)

qplot(x = jitter(age), y = weight, data = nepalData, xlab = "Age in months",  
      ylab = "Weight in kg", ylim = c(0, 12)) +
    geom_line(aes(x = age, y = model6$fitted.values), color = "red", lwd = 2) +
    theme_622()


# Compare linearity
anova(model4, model6)
survey::regTermTest(model6, ~ age_sp1 + age_sp2 + age_sp3)


# Compare models
AIC(model4, model5, model6)

AIC.table <- tibble(
    `Model (# parameters)` = c("Linear (2)", "Monthly means (13)", "Linear spline (5)"),
    `parameters` = c(2, 13, 5),
    `Residual sum of squares (RSS)` = c(anova(model4)$`Sum Sq`[length(anova(model4)$`Sum Sq`)],
                                        anova(model5)$`Sum Sq`[length(anova(model5)$`Sum Sq`)],
                                        anova(model6)$`Sum Sq`[length(anova(model6)$`Sum Sq`)]),
    `Residual mean square (MSE)` = c(anova(model4)$`Mean Sq`[length(anova(model4)$`Mean Sq`)],
                                     anova(model5)$`Mean Sq`[length(anova(model5)$`Mean Sq`)],
                                     anova(model6)$`Mean Sq`[length(anova(model6)$`Mean Sq`)]),
    `AIC.lm` = nrow(nepalData)*(log(2*pi*`Residual sum of squares (RSS)`/nrow(nepalData))+1) +
        2*`parameters`,
    `log-likelihood` = sapply(c(logLik(model4), logLik(model5), logLik(model6)), as.numeric),
    `AIC.glm` = -2*(`log-likelihood`) + 2*`parameters`
) %>%
    mutate(`parameters` = NULL)

AIC.huxtable <- huxtable::as_hux(AIC.table) %>%
    huxtable::insert_row("", "Using the regress command", "", "", "Using the glm command", "") %>%
    huxtable::merge_cells(1, 2:4) %>%
    huxtable::merge_cells(1, 5:6) %>%
    huxtable::set_header_rows(1, TRUE) %>%
    huxtable::style_headers(bold = TRUE)
huxtable::number_format(AIC.huxtable) <- 2
AIC.huxtable
```